---
title: false
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE)
```

I am going to load the entire `ggplot2` library but will otherwise namespace (i.e., package_name::function_name) calls to functions. This is helpful as it allows us to call functions without loading entire libraries. Note, though, that the library must still be installed: in addition to the `ggplot2` library, you will need also the `dplyr`, `tidyr`, and `readr` libraries.

```{r libraries, echo=TRUE}
library(ggplot2)
```

Identify a country code for working with data from a single country.

```{r parameters, echo=TRUE}
country_code <- "US"
```

Read the data from the `data` directory in this repository.

### load-data

```{r load-data, echo=TRUE, message=FALSE}
tb <- readr::read_csv("tb.csv")
```

### tidy the data

```{r tidy-the-data, echo=TRUE}

tb_tidy <- tb |>
  tidyr::pivot_longer(
    cols = c(-iso2, -year),
    names_to = "code",
    values_to = "count"
    ) |>
  tidyr::separate(
    col = code,
    into = c("gender", "age"),
    sep = "_"
    ) |>
  dplyr::filter(!is.na(count))

```


### filter the data

```{r filter-the-data, echo=TRUE}

tb_filtered <- tb_tidy |>
  dplyr::filter(
    iso2 == country_code,
    !age %in% c("u", "014", "04", "514")
  )

```

### plot with fill

```{r plot-with-fill, eval=TRUE}

tb_filtered |>
  ggplot(aes(x = year, y = count, fill = gender)) +
  geom_bar(stat = "identity", position = "fill") +
  facet_grid(~ age) +
  scale_fill_manual(
    "",
    values=c("#486030", "#C03018")
  )

```

### plot sans fill

```{r plot-sans-fill, eval=TRUE}

tb_filtered |>
  ggplot(aes(x = year, y = count, fill = gender)) +
  geom_bar(stat = "identity") +
  facet_grid(~ age) +
  scale_fill_manual(
    "",
    values=c("#486030", "#C03018")
  )

```

### faceted by gender and age

```{r faceted, eval=TRUE}

tb_filtered |>
  ggplot(aes(x = year, y = count, fill = gender)) +
  geom_bar(stat = "identity") +
  facet_grid(gender ~ age) +
  scale_fill_manual(
    "",
    values=c("#486030", "#C03018")
  )

```

### polar

```{r polar, eval=TRUE}

tb_filtered |>
  ggplot(aes(x = year, y = count, fill = gender)) +
  geom_bar(stat = "identity") +
  facet_grid(gender ~ age) +
  scale_fill_manual(
    "",
    values=c("#486030", "#C03018")
    ) +
  coord_polar()

```
