---
title: "tidy_jeopardy"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We can start with a few simple ones, like `mtcars` and `iris`.


### atmospheric conditions

```{r}
sonde <- readr::read_csv("data/CR1000_soil_01_feb272012_jul172013.dat") |>
  head(n = 10)
```


### tuberculosis

```{r}
TB <- readr::read_csv("data/tb.csv") |>
  head(n = 10)
```


### weather

From: [AZMET](https://cals.arizona.edu/azmet/)

```{r}
azmet <- read.csv(
  file = "https://cals.arizona.edu/azmet/data/1202rd.txt",
  header = FALSE,
  sep = ","
)
```

The data appear tidy but it is difficult to tell without knowing the meaning of the variables. We will add the column names to provide some clarity.

```{r}
azmet_cols <- c(
  "Year",
  "Day_of_Year",
  "Station_Number",
  "Air_Temp_Max",
  "Air_Temp_Min",
  "Air_Temp_Mean",
  "RH_Max",
  "RH_Min",
  "RH_Mean",
  "VPD_Mean",
  "Solar_Rad_Total",
  "Precipitation_Total",
  "2-inch_Soil_Temp_Max_(changed_to__4-inch_in_1999)",
  "2-inch_Soil_Temp_Min_(changed_to__4-inch_in_1999)",
  "2-inch_Soil_Temp_Mean_(changed_to__4-inch_in_1999)",
  "4-inch_Soil_Temp_Max_(changed_to_20-inch_in_1999)",
  "4-inch_Soil_Temp_Min_(changed_to_20-inch_in_1999)",
  "4-inch_Soil_Temp_Mean_(changed_to_20-inch_in_1999)",
  "Wind_Speed_Mean",
  "Wind_Vector_Magnitude_for_Day",
  "Wind_Vector_Direction_for_Day",
  "Wind_Direction_Standard_Deviation_for_Day",
  "Max_Wind_Speed",
  "Reference_ET",
  "Heat_Units_(30/12.8_C)_(86/55_F)"
)
```

```{r}
colnames(azmet) <- azmet_cols
```


### NFL player routes

[nfl-player-routes](https://docs.google.com/presentation/d/16oZE5sBX8JTXnWZCuPOx054S6wypQaNqB67BG9gPECI/edit?usp=sharing)


### watersheds

Watersheds near Los Alamos, New Mexico

```{r}
nm_watersheds <- sf::st_read(
  dsn = "data/new_mexico_ws.geojson"
)
```

More context: sampling locations at the outflow of each watershed.

```{r}
nm_pour_points <- sf::st_read(
  dsn = "data/new_mexico_points.geojson"
)
```

Another way to look at these (spatial) data.

```{r}
mapview::mapview(x = nm_watersheds) +
mapview::mapview(x = nm_pour_points)
```
